High-level cognition during story listening is reflected in high-order dynamic correlations in neural activity patterns

### 2020-11-13 paper/old.tex
minor tweaks made after going over response letter again, also added track changes doc

\begin{abstract}
  Our thoughts arise from coordinated patterns of interactions between
  brain structures that change with our ongoing experiences.
  High-order dynamic correlations in neural activity patterns reflect
  different subgraphs of the brain's functional connectome that display
  homologous lower-level dynamic correlations.  We tested the
  hypothesis that high-level cognition is reflected in high-order
  dynamic correlations in brain activity patterns.  We developed an
  approach to estimating high-order dynamic correlations in timeseries
  data, and we applied the approach to neuroimaging data collected as
  human participants either listened to a ten-minute story or listened
  to a temporally scrambled version of the story.  We trained
  across-participant pattern classifiers to decode (in held-out data)
  when in the session each neural activity snapshot was collected.  We
  found that classifiers trained to decode from high-order dynamic
  correlations yielded the best performance on data collected as
  participants listened to the (unscrambled) story.  By contrast,
  classifiers trained to decode data from scrambled versions of the
  story yielded the best performance when they were trained using
  first-order dynamic correlations or non-correlational activity
  patterns.  We suggest that as our thoughts become more complex, they
  are reflected in higher-order patterns of dynamic network
  interactions throughout the brain.
\end{abstract}

@@

\section*{Introduction}
A central goal in cognitive neuroscience is to elucidate the
\textit{neural code}: the mapping between (a) mental states or
cognitive representations and (b) neural activity patterns. One means
of testing models of the neural code is to ask how accurately that
model is able to ``translate'' neural activity patterns into known (or
hypothesized) mental states or cognitive
representations~\citep[e.g.,][]{HaxbEtal01, NormEtal06, TongPrat12,
  MitcEtal08a, KamiTong05, NishEtal11, PereEtal18, HuthEtal12,
  HuthEtal16}.  Training decoding models on different types of neural
features (Fig.~\ref{fig:patterns}a) can also help to elucidate which
specific aspects of neural activity patterns are informative about
cognition-- and, by extension, which types of neural activity patterns
might compose the neural code.  For example, prior work has used
region of interest analyses to estimate the anatomical locations of
specific neural representations~\citep[e.g.,][]{EtzeEtal09}, or to
compare the relative contributions to the neural code of multivariate
activity patterns versus dynamic correlations between neural activity
patterns~\citep[e.g.,][]{MannEtal18, FongEtal19}.  An emerging theme
in this literature is that cognition is mediated by dynamic
interactions between brain structures~\citep{Gros88, Fris00, SporHone06, BassEtal06,
  Turk13, DemeEtal19, SoloEtal19, LuriEtal18, PretEtal17, ZouEtal19,
  MackEtal17}.

@@

Studies of the neural code to date have primarily focused on
univariate or multivariate neural patterns~\citep[for review
see][]{NormEtal06}, or (more recently) on patterns of dynamic
first-order correlations~\citep[i.e., interactions between pairs of
brain structures;][]{MannEtal18, FongEtal19, LuriEtal18, PretEtal17,
  ZouEtal19, DemeEtal19}.  What might the future of this line of work
hold?  For example, is the neural code implemented through
higher-order interactions between brain structures~\citep[e.g.,
see][]{ReimEtal17}?  Second-order correlations reflect
\textit{homologous} patterns of correlation.  In other words, if the
dynamic patterns of correlations between two regions, $A$ and $B$, are
similar to those between two other regions, $C$ and $D$, this would be
reflected in the second-order correlations between ($A$--$B$) and
($C$--$D$).  In this way, second-order correlations identify
similarities and differences between subgraphs of the brain's
connectome.  Analogously, third-order correlations reflect homologies
between second-order correlations-- i.e., homologous patterns of
homologous interactions between brain regions.  More generally,
higher-order correlations reflect homologies between patterns of
lower-order correlations.  We can then ask: which ``orders'' of
interaction are most reflective of high-level cognitive processes?

One reason one might expect to see homologous networks in a dataset is
related to the notion that network dynamics reflect ongoing neural
computations or cognitive processing~\citep[e.g.,][]{BeatEtal16}.  If
the nodes in two brain networks are interacting (within each network)
in similar ways then, according to our characterization of network
dynamics, we refer to the similarities between those patterns of
interaction as higher-order correlations.  When higher-order
correlations are themselves changing over time, we can also attempt to
capture and characterize those high-order dynamics.

@@

We validated our approach using synthetic data where the underlying
correlations were known.  We then applied our framework to a
neuroimaging dataset collected as participants listened to either an
audio recording of a ten-minute story, listened to a temporally
scrambled version of the story, or underwent a resting state
scan~\citep{SimoEtal16}.  Temporal scrambling has been used in a
growing number of studies, largely by Uri Hasson's group, to identify
brain regions that are sensitive to higher-order and longer-timescale
information (e.g., cross-sensory integration, rich narrative meaning,
complex situations, etc.) versus regions that are primarily sensitive
to low-order (e.g., sensory) information.  For example,
\cite{HassEtal08} argues that when brain areas are sensitive to fine
versus coarse temporal scrambling, this indicates that they are
``higher order'' in the sense that they process contextual information
pertaining to further-away timescales.  By contrast, low-level
regions, such as primary sensory cortices, do not meaningfully change
their responses (after correcting for presentation order) even when
the stimulus is scrambled at fine timescales.

We used a subset of the story listening and rest data to train
across-participant classifiers to decode listening times (of groups of
participants) using a blend of neural features (comprising neural
activity patterns, as well as different orders of dynamic correlations
between those patterns that were inferred using our computational
framework).  We found that both the PCA-based and eigenvector
centrality-based approaches yielded neural patterns that could be used
to decode accurately (i.e., well above chance).  Both approaches also
yielded the best decoding accuracy for data collected during (intact)
story listening when high-order (PCA: second-order; eigenvector
centrality: fourth-order) dynamic correlation patterns were included
as features.  When we trained classifiers on the scrambled stories or
resting state data, only (relatively) lower-order dynamic patterns
were informative to the decoders.  Taken together, our results
indicate that high-level cognition is supported by high-order dynamic
patterns of communication between brain structures.

\section*{Results}
We sought to understand whether high-level cognition is reflected in
dynamic patterns of high-order correlations.  To that end, we
developed a computational framework for estimating the dynamics of
stimulus-driven high-order correlations in multivariate timeseries
data (see \textit{Dynamic inter-subject functional connectivity
  (DISFC)} and \textit{Dynamic higher-order correlations}).  We
evaluated the efficacy of this framework at recovering known patterns
in several synthetic datasets (see \textit{Synthetic data: simulating
  dynamic first-order correlations} and \textit{Synthetic data:
  simulating dynamic higher-order correlations}).  We then
applied the framework to a public fMRI dataset collected as
participants listened to an auditorily presented story, listened to a
temporally scrambled version of the story, or underwent a resting
state scan (see \textit{Functional neuroimaging data collected during
  story listening}).  We used the relative decoding accuracies of
classifiers trained on different sets of neural features to estimate
which types of features reflected ongoing cognitive processing.

\subsection*{Recovering known dynamic correlations from synthetic
  data}
\subsubsection*{Recovering dynamic first-order correlations}
We generated synthetic datasets that differed in how the underlying first-order
correlations changed over time.  For each dataset, we applied
Equation~\ref{eqn:timecorr} with a variety of kernel shapes and
widths.  We assessed how well the true underlying correlations at each
timepoint matched the recovered correlations
(Fig.~\ref{fig:synthetic}).  For every kernel and dataset we tested,
our approach recovered the correlation dynamics we embedded into the
data.  However, the quality of these recoveries varied across
different synthetic datasets in a kernel-dependent way.

\begin{figure}[tp]
  \centering
  \includegraphics[width=\textwidth]{figs/synthetic_data}
  \caption{\textbf{Recovering known dynamic first-order correlations
      from synthetic data.}  Each panel displays the average
    correlations between the vectorized upper triangles of the
    recovered correlation matrix at each timepoint and either the true
    underlying correlation at each timepoint or a reference
    correlation matrix.  (The averages are taken across 10 different
    randomly generated synthetic datasets of the given category.)
    Error ribbons denote 95\% confidence intervals (taken across
    datasets). Different colors denote different kernel shapes, and
    the shading within each color family denotes the kernel width
    parameter.  For a complete description of each synthetic dataset,
    see \textit{Synthetic data: simulating dynamic first-order
      correlations}.  \textbf{a. Constant correlations.}  These
    datasets have a stable (unchanging) underlying correlation matrix.
    \textbf{b. Random correlations.}  These datasets are generated
    using a new independently drawn correlation matrix at each new
    timepoint.  \textbf{c. Ramping correlations.}  These datasets are
    generated by smoothly varying the underlying correlations between
    the randomly drawn correlation matrices at the first and last
    timepoints.  The left panel displays the correlations between the
    recovered dynamic correlations and the underlying ground truth
    correlations.  The middle panel compares the recovered
    correlations with the \textit{first} timepoint's correlation
    matrix.  The right panel compares the recovered correlations with
    the \textit{last} timepoint's correlation matrix.
    \textbf{d. Event-based correlations.}  These datasets are each
    generated using five randomly drawn correlation matrices that each
    remain stable for a fifth of the total timecourse.  The left panel
    displays the correlations between the recovered dynamic
    correlations and the underlying ground truth correlations.  The
    right panels compare the recovered correlations with the
    correlation matrices unique to each event.  The vertical lines
    denote event boundaries.}
  \label{fig:synthetic}
\end{figure}

\subsubsection*{Recovering dynamic higher-order correlations}
Following our approach to evaluating our ability to recover known
dynamic first-order correlations from synthetic data, we generated an
analogous second set of synthetic datasets that we designed to exhibit
known dynamic first-order \textit{and} second-order correlations (see
\textit{Synthetic data: simulating dynamic higher-order
  correlations}).  We generated a total of 40 datasets that varied in
how the first-order and second-order correlations changed over time.
We then repeatedly applied Equation~\ref{eqn:timecorr} using the
overall best-performing kernel from our first-order tests (a Laplace
kernel with a width of 20; Fig.~\ref{fig:synthetic}) to assess how
closely the recovered dynamic correlations matched the dynamic
correlations we had embedded into the datasets.

Overall, we found that we could reliably recover both first-order and
second-order correlations from the synthetic data
(Fig.~\ref{fig:higher_order_sims}).  When the correlations were stable
for longer intervals, or changed gradually (constant, ramping, and
event datasets), recovery performance was relatively high, and we were
better able to recover dynamic first-order correlations than
second-order correlations.  We expected that this would happen, given
that errors in our estimation procedure at lower orders necessarily
propagate to higher orders (since lower-order correlations are used to
estimate higher-order correlations).  Interestingly, we also found
that when the correlations were particularly \textit{un}stable (random
datasets), we better recovered second-order correlations.

\begin{figure}[tp]
  \centering
  \includegraphics[width=\textwidth]{figs/higher_order_sims}
  \caption{\textbf{Recovery of simulated first-order and second-order dynamic correlations.}  Each panel displays the average correlations
    between the vectorized upper triangles of the recovered
    correlation matrix and the simulated correlation matrix at each
    timepoint and for each synthetic dataset.  (The averages are taken across 10 different
    randomly generated synthetic datasets of the given category.)
    Error ribbons denote 95\% confidence intervals (taken across
    datasets).  For a complete description of each synthetic dataset,
    see \textit{Synthetic data: simulating dynamic higher-order
      correlations}.  All results in this figure used a Laplace kernel
    (width~$=20$).  \textbf{a. Constant correlations.}
    These first and second order dynamic correlation were
    simulated from datasets have a stable (unchanging) underlying correlation
    matrix.  \textbf{b. Random correlations.}  These first and second order dynamic correlation were
    simulated from datasets are
    generated using a new independently drawn correlation matrix at
    each new timepoint.  \textbf{c. Ramping correlations.}  These first and second order dynamic correlation were
    simulated from
    datasets are generated by smoothly varying the underlying
    correlations between the randomly drawn correlation matrices at
    the first and last timepoints and displays the
    correlations between the recovered dynamic correlations and the
  simulated correlation first and second order correlation matrices.   \textbf{d. Event-based correlations.}  These first and second order dynamic correlation were
    simulated from datasets are
    each generated using five randomly drawn correlation matrices that
    each remain stable for a fifth of the total timecourse.  The
    vertical lines denote event boundaries.}
  \label{fig:higher_order_sims}
\end{figure}

Taken together, our explorations using synthetic data indicated that
we are able to partially, but not perfectly, recover ground truth
dynamic first-order and second-order correlations.  This suggests that
our modeling approach provides a meaningful (if noisy) estimate of
high-order correlations.  We next turned to analyses of human fMRI data
to examine whether the recovered dynamics might reflect the dynamics
of human cognition during a naturalistic story-listening task.

\subsection*{Cognitively relevant dynamic high-order correlations in
  fMRI data}
We used across-participant temporal decoders to identify cognitively
relevant neural patterns in fMRI data (see \textit{Forward inference
  and decoding accuracy}).  The dataset we examined~\citep[collected
by][]{SimoEtal16} comprised four experimental conditions that exposed
participants to stimuli that varied systematically in how cognitively
engaging they were.  The \textit{intact} experimental condition had
participants listen to an audio recording of a 10-minute story.  The
\textit{paragraph}-scrambled experimental condition had participants
listen to a temporally scrambled version of the story, where the
paragraphs occurred out of order (but where the same total set of
paragraphs were presented over the full listening interval).  All
participants in this condition experienced the scrambled paragraphs in
the same order.  The \textit{word}-scrambled experimental condition
had participants listen to a temporally scrambled version of the story
where the words in the story occurred in a random order.  All
participants in the word condition experienced the scrambled words in
the same order.  Finally, in a \textit{rest} experimental condition,
participants lay in the scanner with no overt stimulus, with their
eyes open (blinking as needed).  This dataset provided a convenient
means of testing our hypothesis that different levels of cognitive
processing and engagement are reflected in different orders of brain
activity dynamics.

\begin{figure}[tp]
  \centering
  \includegraphics[width=\textwidth]{figs/decode_level}
  \caption{\textbf{Across-participant timepoint decoding accuracy varies with
      correlation order and cognitive engagement.}
    \textbf{a.~Decoding accuracy as a function of order: PCA.}
    \textit{Order} ($x$-axis) refers to the maximum order of dynamic
    correlations that were available to the classifiers (see
    \textit{Feature weighting and testing}).  The reported
    across-participant decoding accuracies are averaged over all
    kernel shapes and widths (see \textit{Identifying robust decoding
      results}).  The $y$-values are displayed relative to chance
    accuracy (intact: $\frac{1}{300}$; paragraph: $\frac{1}{272}$;
    word: $\frac{1}{300}$; rest: $\frac{1}{400}$).  The error ribbons
    denote 95\% confidence intervals across cross-validation folds
    (i.e., random assignments of participants to the training and test
    sets).  The colors denote the experimental condition.  Arrows
    denote sets of features that yielded reliably higher (upwards
    facing) or lower (downward facing) decoding accuracy than the mean
    of all other features (via a two-tailed test, thresholded at
    $p < 0.05$).  Figure~\ref{fig:ttests} displays additional
    comparisons between the decoding accuracies achieved using
    different sets of neural features.  The circled values represent
    the maximum decoding accuracy within each experimental condition.
    \textbf{b.~Normalized timepoint decoding accuracy as a function of order:
      PCA.}  This panel displays the same results as Panel a, but here
    each curve has been normalized to have a maximum value of 1 and a
    minimum value of 0 (including the upper and lower bounds of the
    respective 95\% confidence intervals).  Panels a and b used PCA to
    project each high-dimensional pattern of dynamic correlations onto
    a lower-dimensional space.  \textbf{c.~Timepoint decoding accuracy as a
      function of order: eigenvector centrality.} This panel is in the
    same format as Panel a, but here eigenvector centrality has been
    used to project the high-dimensional patterns of dynamic
    correlations onto a lower-dimensional space.
    \textbf{d.~Normalized timepoint decoding accuracy as a function of order:
      eigenvector centrality.} This panel is in the same format as
    Panel b, but here eigenvector centrality has been used to project
    the high-dimensional patterns of dynamic correlations onto a
    lower-dimensional space.}
  \label{fig:decoding}
\end{figure}

In brief, we computed timeseries of dynamic high-order correlations
that were similar across participants in each of two randomly assigned
groups: a training group and a test group.  We then trained
classifiers on the training group's data to match each sample from the
test group with a stimulus timepoint.  Each classifier comprised a
weighted blend of neural patterns that reflected up to
$n^\mathrm{th}$-order dynamic correlations (see \textit{Feature
  weighting and testing}, Fig.~\ref{fig:pipeline}).  We repeated this process for
$n \in \left\{ 0, 1, 2, ..., 10 \right\}$.  Our examinations of
synthetic data suggested that none of the kernels we examined were
``universal'' in the sense of optimally recovering underlying
correlations regardless of the temporal structure of those
correlations.  We found a similar pattern in the (real) fMRI data,
whereby different kernels yielded different decoding accuracies, but
no single kernel emerged as the clear ``best.''  In our analyses of
neural data, we therefore averaged our decoding results over a variety
of kernel shapes and widths in order to identify results that were
robust to specific kernel parameters (see \textit{Identifying robust
  decoding results}).

Our approach to estimating dynamic high-order correlations entails
mapping the high-dimensional feature space of correlations
(represented by a $T$ by
$\mathcal{O}(K^2)$ matrix) onto a lower-dimensional feature space
(represented by a $T$ by $K$ matrix).
We carried out two sets of analyses that differed in how this mapping
was computed.  The first set of analyses used PCA to find a
low-dimensional embedding of the original dynamic correlation matrices
(Fig.~\ref{fig:decoding}a,b).  The second set of analyses
characterized correlations in dynamics of each feature's eigenvector
centrality, but did not preserve the underlying activity dynamics
(Fig.~\ref{fig:decoding}c,d).

\begin{figure}[tp]
\centering
\includegraphics[width=1\textwidth]{figs/stats_heatmaps}
\caption{\textbf{Statistical summary of decoding accuracies for
    different neural features.}  Each column of matrices displays decoding
  results for one experimental condition (intact, paragraph, word, and
  rest).  We considered dynamic activity patterns (order 0) and
  dynamic correlations at different orders (order $> 0$).  We used
  two-tailed $t$-tests to compare the
  distributions of decoding accuracies obtained using each pair of
  features.  The distributions for each feature reflect the set of
  average decoding accuracies (across all kernel parameters), obtained
  for each random assignment of training and test groups.
  In the upper triangles of each matrix, warmer colors (positive $t$-values) indicate that the
  neural feature indicated in the given row yielded higher accuracy than the
  feature indicated in the given column.  Cooler colors (negative
  $t$-values) indicate that
  the feature in the given row yielded lower decoding accuracy than
  the feature in the given column.  The lower triangles of each map
  denote the corresponding $p$-values for the $t$-tests.  The diagonal
  entries display the relative average optimized weight given to each type of feature, in
  a decoder that included all feature types (see \textit{Feature
    weighting and testing}).}
\label{fig:ttests}
\end{figure}

\begin{figure}[tp]
  \centering
  \includegraphics[width=\textwidth]{figs/most_abs}
  \caption{\textbf{Top terms associated with the endpoints of the
      strongest correlations.}  Each color corresponds to one order of
    inter-subject functional correlations.  To calculate the dynamic
    correlations, eigenvector centrality has been used to project
    the high-dimensional patterns of dynamic correlations onto a
    lower-dimensional space at each previous order, which allows us
    the map the brain regions at each order by retaining the
    features of the original space. The inflated brain plots
    display the locations of the endpoints of the 10 strongest
    (absolute value) correlations at each order, thresholded at 0.999,
    and projected onto the cortical surface~\citep{CombEtal19}.  The
    lists of terms on the right display the top five Neurosynth
    terms~\citep{RubiEtal17} decoded from the corresponding brain maps
    for each order.  Each row displays data from a different
    experimental condition.  Additional maps and their corresponding
    Neurosynth terms may be found in the \textit{Supplementary
      materials} (intact: Fig.~\intact; paragraph: Fig.~\para; word:
    Fig.~\word; rest: Fig.~\rest).}
  \label{fig:neurosynth}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Discussion}
We tested the hypothesis that high-level cognition is reflected in
high-order brain network dynamics~\citep[e.g., see][]{SoloEtal19,
  ReimEtal17}.  We examined high-order network dynamics in functional
neuroimaging data collected during a story listening experiment.  When
participants listened to an auditory recording of the story,
participants exhibited similar high-order brain network dynamics.  By
contrast, when participants instead listened to temporally scrambled
recordings of the story, only lower-order brain network dynamics were
similar across participants.  Our results indicate that higher orders
of network interactions support higher-level aspects of cognitive
processing (Fig.~\ref{fig:discussion}).

\begin{figure}[tp]
  \centering
  \includegraphics[width=0.6\textwidth]{figs/discussion}
  \caption{\textbf{Proposed high-order network dynamics underlying
      high-level cognition during story listening.}  Schematic depicts
    higher orders of
    network interactions supporting higher-level aspects of cognitive
    processing.  When tasks evoke richer, deeper, and/or higher-level
    processing, this is reflected in higher-order network
    interactions.}
  \label{fig:discussion}
\end{figure}

The notion that cognition is reflected in (and possibly mediated by)
patterns of first-order network dynamics has been suggested by or
proposed in myriad empirical studies and
reviews~\citep[e.g.,][]{DemeEtal19, Turk13, LuriEtal18, FongEtal19,
  ParkEtal18b, PretEtal17, MannEtal18, RoyEtal19, LiegEtal19, ZouEtal19,
  ChanGlov10, GonzEtal19}.  Our study extends this line of work by
finding cognitively relevant \textit{higher-order} network dynamics
that reflect ongoing cognition.  Our findings complement other work
that uses graph theory and topology to characterize how brain networks
reconfigure during cognition~\citep[e.g.,][]{BassEtal06, ZhenEtal19,
  McInJirs19, TokeSomm19, SizeEtal18, ReimEtal17, BetzEtal19}.

An open question not addressed by our study pertains to how different
structures integrate incoming information with different time
constants.  For example, one line of work suggests that the cortical
surface comprises a structured map such that nearby brain structures
process incoming information at similar timescales.  Low-level sensory
areas integrate information relatively quickly, whereas higher-level
regions integrate information relatively slowly~\citep{BaldEtal17,
  HassEtal08, HassEtal15, HoneEtal12a, LernEtal11, LernEtal14,
  ChieHone19}.  A similar hierarchy appears to play a role in
predicting future events~\citep{LeeEtal20}.  Other
related work in human and mouse brains indicates that the temporal
response profile of a given brain structure may relate to how strongly
connected that structure is with other brain
areas~\citep{FallEtal19}.  Further study is needed to understand the
role of temporal integration at different scales of network
interaction, and across different anatomical structures. Importantly,
our analyses do not speak to the physiological basis of higher-order
dynamics, and could reflect nonlinearities, chaotic patterns,
non-stationarities, or multistability, etc. However, our decoding analyses do indicate
that higher-order dynamics are
consistent across individuals, and therefore unlikely to be driven by
non-stimulus-driven dynamics which are unlikely to be similar across
individuals.

Identifying high-order network dynamics associated with high-level
cognition required several important methods advances.  First, we used
kernel-based dynamic correlations to extended the notion of (static)
inter-subject functional connectivity~\citep{SimoEtal16} to a dynamic
measure of inter-subject functional connectivity (DISFC) that does not
rely on sliding windows~\citep[e.g., as in][]{MannEtal18}, and that
may be computed at individual timepoints.  This allowed us to
precisely characterize stimulus-evoked network dynamics that were
similar across individuals.  Second, we developed a computational
framework for efficiently and scalably estimating high-order dynamic
correlations.  Our approach uses dimensionality reduction algorithms
and graph measures to obtain low-dimensional embeddings of patterns of
network dynamics.  Third, we developed an analysis framework for
identifying robust decoding results by carrying out our analyses using
a range of parameter values and then identifying which results were
robust to specific parameter choices.

\subsection*{Concluding remarks}
Methods for characterizing the dynamics of high-order correlations in
neural data provide a window into the neural basis of cognition.  By
showing that high-level cognition is reflected in high-order network
dynamics, we have elucidated the next step on the path towards
understanding the neural basis of cognition.

\section*{Methods}
Our general approach to efficiently estimating high-order dynamic
correlations comprises four general steps (Fig.~\ref{fig:methods}).
First, we derive a kernel-based approach to computing dynamic pairwise
correlations in a $T$ (timepoints) by $K$ (features) multivariate
timeseries, $\mathbf{X}_0$.  This yields a $T$ by $\mathcal{O}(K^2)$
matrix of dynamic correlations, $\mathbf{Y}_1$, where each row
comprises the upper triangle and diagonal of the correlation matrix at a single
timepoint, reshaped into a row vector (this reshaped vector is
$\left( \frac{K^2 - K}{2} + K \right)$-dimensional).  Second, we apply a dimensionality
reduction step to project the matrix of dynamic correlations back onto
a $K$-dimensional space.  This yields a $T$ by $K$ matrix,
$\mathbf{X}_1$, that reflects an approximation of the dynamic
correlations reflected in the original data.  Third, we use repeated
applications of the kernel-based dynamic correlation step to
$\mathbf{X}_n$ and the dimensionality reduction step to the resulting
$\mathbf{Y}_{n+1}$ to estimate high-order dynamic correlations.  Each
application of these steps to a $T$ by $K$ time series $\mathbf{X}_n$
yields a $T$ by $K$ matrix, $\mathbf{X}_{n+1}$, that reflects the
dynamic correlations between the columns of $\mathbf{X}_n$.  In this
way, we refer to $n$ as the \textit{order} of the timeseries, where
$\mathbf{X}_0$ (order 0) denotes the original data and $\mathbf{X}_n$
denotes (approximated) $n^\mathrm{th}$-order dynamic correlations
between the columns of $\mathbf{X}_0$.  Finally, we use a
cross-validation--based decoding approach to evaluate how well
information contained in a given order (or weighted mixture of orders)
may be used to decode relevant cognitive states.  If including a given
$\mathbf{X}_n$ in the feature set yields higher classification
accuracy on held-out data, we interpret this as evidence that the
given cognitive states are reflected in patterns of
$n^\mathrm{th}$-order correlations.

All of the code used to produce
the figures and results in this manuscript, along with links to the
corresponding datasets, may be found at
\href{https://github.com/ContextLab/timecorr-paper}{github.com/ContextLab/timecorr-paper}.
In addition, we have released a Python toolbox for computing dynamic
high-order correlations in timeseries data; our toolbox may be found
at
\href{https://timecorr.readthedocs.io/}{timecorr.readthedocs.io}.

@@

  \paragraph*{Ramping.}  We generated a third set of 10 synthetic
  datasets whose underlying correlations changed gradually over time.
  For each dataset, we constructed two \textit{anchor} covariance
  matrices using Equation~\ref{eqn:randcorr},
  $\mathbf{\Sigma}_{\mathrm{start}}$ and
  $\mathbf{\Sigma}_{\mathrm{end}}$.  For each of the 300 timepoints in
  each dataset, we drew the observations from a multivariate Gaussian
  distribution whose covariance matrix at each timepoint $t \in
  \left[0, ..., 299\right]$ was given by
  \begin{align}
    \mathbf{\Sigma}_t = \left( 1 - \frac{t}{299} \right)
    \mathbf{\Sigma}_{\mathrm{start}} + \frac{t}{299}\mathbf{\Sigma}_{\mathrm{end}}.
  \end{align}
The gradually changing correlations underlying these datasets allow us
to evaluate the recovery of dynamic correlations when each timepoint's
correlation matrix is unique (as in the random datasets), but where
the correlation dynamics are structured and exhibit first-order
autocorrelations (as in the constant datasets).

>>>>

\begin{abstract}
  Our thoughts arise from coordinated patterns of interactions between
  brain structures that change with our ongoing experiences.
  High-order dynamic correlations in neural activity patterns reflect
  different subgraphs of the brain's connectome that display
  homologous lower-level dynamic correlations.  We tested the
  hypothesis that high-level cognition is supported by high-order
  dynamic correlations in brain activity patterns.  We developed an
  approach to estimating high-order dynamic correlations in timeseries
  data, and we applied the approach to neuroimaging data collected as
  human participants either listened to a ten-minute story, listened
  to a temporally scrambled version of the story, or underwent a
  resting state scan.  We trained across-participant pattern
  classifiers to decode (in held-out data) when in the session each
  neural activity snapshot was collected.  We found that classifiers
  trained to decode from high-order dynamic correlations yielded the
  best performance on data collected as participants listened to the
  (unscrambled) story.  By contrast, classifiers trained to decode
  data from scrambled versions of the story or during the resting
  state scan yielded the best performance when they were trained using
  first-order dynamic correlations or non-correlational activity patterns.  We
  suggest that as our thoughts become more complex, they are supported
  by higher-order patterns of dynamic network interactions throughout
  the brain.
\end{abstract}

@@

\section*{Introduction}
A central goal in cognitive neuroscience is to elucidate the
\textit{neural code}: the mapping between (a) mental states or
cognitive representations and (b) neural activity patterns. One means
of testing models of the neural code is to ask how accurately that
model is able to ``translate'' neural activity patterns into known (or
hypothesized) mental states or cognitive
representations~\citep[e.g.,][]{HaxbEtal01, NormEtal06, TongPrat12,
  MitcEtal08a, KamiTong05, NishEtal11, PereEtal18, HuthEtal12,
  HuthEtal16}.  Training decoding models on different types of neural
features (Fig.~\ref{fig:patterns}a) can also help to elucidate which
specific aspects of neural activity patterns are informative about
cognition-- and, by extension, which types of neural activity patterns
might comprise the neural code.  For example, prior work has used
region of interest analyses to estimate the anatomical locations of
specific neural representations~\citep[e.g.,][]{EtzeEtal09}, or to
compare the relative contributions to the neural code of multivariate
activity patterns versus dynamic correlations between neural activity
patterns~\citep[e.g.,][]{MannEtal18, FongEtal19}.  An emerging theme
in this literature is that cognition is mediated by dynamic
interactions between brain structures~\citep{SporHone06, BassEtal06,
  Turk13, DemeEtal19, SoloEtal19, LuriEtal18, PretEtal17, ZouEtal19,
  MackEtal17}.

@@

Studies of the neural code to date have primarily focused on
univariate or multivariate neural patterns~\citep[for review
see][]{NormEtal06}, or (more recently) on patterns of dynamic
first-order correlations~\citep[i.e., interactions between pairs of
brain structures;][]{MannEtal18, FongEtal19, LuriEtal18, PretEtal17,
  ZouEtal19, DemeEtal19}.  We wondered what the future of this line of
work might hold.  For example, is the neural code mediated by
higher-order interactions between brain structures~\citep[e.g.,
see][]{ReimEtal17}?  Second-order correlations reflect
\textit{homologous} patterns of correlation.  In other words, if the
dynamic patterns of correlations between two regions, $A$ and $B$, are
similar to those between two other regions, $C$ and $D$, this would be
reflected in the second-order correlations between ($A$--$B$) and
($C$--$D$).  In this way, second-order correlations identify
similarities and differences between subgraphs of the brain's
connectome.  Analogously, third-order correlations reflect homologies
between second-order correlations-- i.e., homologous patterns of
homologous interactions between brain regions.  More generally,
higher-order correlations reflect homologies between patterns of
lower-order correlations.  We can then ask: which ``orders'' of
interaction are most reflective of high-level cognitive processes?

@@

We validated our approach using synthetic data where the underlying
correlations were known.  We then applied our framework to a
neuroimaging dataset collected as participants listened to either an
audio recording of a ten-minute story, listened to a temporally scrambled
version of the story, or underwent a resting state
scan~\citep{SimoEtal16}.  We used a subset of the data to train
across-participant classifiers to decode listening times using a blend
of neural features (comprising neural activity patterns, as well as
different orders of dynamic correlations between those patterns that
were inferred using our computational framework).  We found that both
the PCA-based and eigenvector centrality-based approaches yielded
neural patterns that could be used to decode accurately (i.e., well above
chance).  Both approaches also yielded the best decoding accuracy for
data collected during (intact) story listening when high-order (PCA:
second-order; eigenvector centrality: fourth-order) dynamic
correlation patterns were included as features.  When we trained
classifiers on the scrambled stories or resting state data, only
lower-order dynamic patterns were informative to the decoders.  Taken
together, our results indicate that high-level cognition is supported
by high-order dynamic patterns of communication between brain
structures.

\section*{Methods}
Our general approach to efficiently estimating high-order dynamic
correlations comprises four general steps (Fig.~\ref{fig:methods}).
First, we derive a kernel-based approach to computing dynamic pairwise
correlations in a $T$ (timepoints) by $K$ (features) multivariate
timeseries, $\mathbf{X}_0$.  This yields a $T$ by $\mathcal{O}(K^2)$
matrix of dynamic correlations, $\mathbf{Y}_1$, where each row
comprises the upper triangle and diagonal of the correlation matrix at a single
timepoint, reshaped into a row vector (this reshaped vector is
$\left( \frac{K^2 - K}{2} + K \right)$-dimensional).  Second, we apply a dimensionality
reduction step to project the matrix of dynamic correlations back onto
a $K$-dimensional space.  This yields a $T$ by $K$ matrix,
$\mathbf{X}_1$, that reflects an approximation of the dynamic
correlations reflected in the original data.  Third, we use repeated
applications of the kernel-based dynamic correlation step to
$\mathbf{X}_n$ and the dimensionality reduction step to the resulting
$\mathbf{Y}_{n+1}$ to estimate high-order dynamic correlations.  Each
application of these steps to a $T$ by $K$ time series $\mathbf{X}_n$
yields a $T$ by $K$ matrix, $\mathbf{X}_{n+1}$, that reflects the
dynamic correlations between the columns of $\mathbf{X}_n$.  In this
way, we refer to $n$ as the \textit{order} of the timeseries, where
$\mathbf{X}_0$ (order 0) denotes the original data and $\mathbf{X}_n$
denotes (approximated) $n^\mathrm{th}$-order dynamic correlations
between the columns of $\mathbf{X}_0$.  Finally, we use a
cross-validation--based decoding approach to evaluate how well
information contained in a given order (or weighted mixture of orders)
may be used to decode relevant cognitive states.  If including a given
$\mathbf{X}_n$ in the feature set yields higher classification
accuracy on held-out data, we interpret this as evidence that the
given cognitive states are reflected in patterns of
$n^\mathrm{th}$-order correlations.  All of the code used to produce
the figures and results in this manuscript, along with links to the
corresponding datasets, may be found at
\href{https://github.com/ContextLab/timecorr-paper}{github.com/ContextLab/timecorr-paper}.
In addition, we have released a Python toolbox for computing dynamic
high-order correlations in timeseries data; our toolbox may be found
at
\href{https://timecorr.readthedocs.io/}{timecorr.readthedocs.io}.

@@

  \paragraph*{Ramping.}  We generated a third set of 10 synthetic
  datasets whose underlying correlations changed gradually over time.
  For each dataset, we constructed two \textit{anchor} covariance
  matrices using Equation~\ref{eqn:randcorr},
  $\mathbf{\Sigma}_{\mathrm{start}}$ and
  $\mathbf{\Sigma}_{\mathrm{end}}$.  For each of the 300 timepoints in
  each dataset, we drew the observations from a multivariate Gaussian
  distribution whose covariance matrix at each timepoint $t \in
  \left[0, ..., 299\right]$ was given by
  \begin{align}
    \mathbf{\Sigma}_t = \left( 1 - \frac{t}{299} \right)
    \mathbf{\Sigma}_{\mathrm{start}} + \frac{t}{299}\mathbf{\Sigma}_{\mathrm{end}}.
  \end{align}
The gradually changing correlations underlying these datasets allow us
to evaluate the recovery of dynamic correlations when each timepoint's
correlation matrix is unique (as in the random datasets), but where
the correlation dynamics are structured.

\paragraph*{Event.} We generated a fourth set of 10 synthetic datasets
whose underlying correlation matrices exhibited prolonged intervals of
stability, interspersed with abrupt changes.  For each dataset, we
used Equation~\ref{eqn:randcorr} to generate 5 random covariance
matrices.  We constructed a timeseries where each set of 60 consecutive samples
was drawn from a Gaussian with the same covariance matrix.  These
datasets were intended to simulate a system that undergoes occasional
abrupt state changes.
